### 2024-09-20 src/Proposal.Rnw
Update

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsfonts}
\usepackage{attachfile}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{geometry}
\usepackage{wrapfig}
\usepackage[colorinlistoftodos]{todonotes}
% For including and formatting code
\usepackage{listings}
\usepackage{xcolor}

% Define global constants
\setlength{\marginparwidth}{2cm} % Adjusting marginparwidth for todonotes
\setlength{\intextsep}{5pt} % Vertical space above & below [h] floats
\setlength{\columnsep}{10pt} % Space between columns
\setlength{\abovecaptionskip}{0pt} % Space above caption
\setlength{\belowcaptionskip}{10pt} % Space below caption
\geometry{left=1.5cm, right=1.5cm, top=2cm, bottom=2cm} % Set the page to landscape with custom margins
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}  % Defines a new command for the horizontal lines, change thickness here
\newcommand{\SubSec}[1]{\vspace{1em}\noindent\textsc{\textbf{#1}}\vspace{0.5em} \\} % Custom command to mimic subsections
\newcommand{\Answer}{\vspace{0.2cm}\textbf{Answer:}\vspace{0.2cm}}

% R Libraries import
<<setup, include=FALSE>>=
knitr::opts_chunk$set(fig.path='assets/')
@

\begin{document}

%----------------------------------------------------------------------------------------
% Title Page
%----------------------------------------------------------------------------------------
\begin{titlepage}
\center{}
\textsc{\LARGE The University of Newcastle}\\[0.5cm]
\textsc{\large College of Engineering, Science and Environment}\\[0.5cm]
\textsc{\large School of Information and Physical Sciences}\\[0.5cm]
\includegraphics[scale=.3]{assets/_uni_logo.png}\\[0.5cm]
\textsc{\Large Statistical Inference}\\[0.3cm]
\textsc{\Large STAT2300: Semester 2, 2024}\\[0.3cm]

% TITLE BAR

\HRule{} \\[0.4cm]
{\huge \bfseries Group Project - Proposal } \\ [0.4cm]
\HRule{} \\[1cm]

% AUTHOR SECTION

\begin{minipage}{1.0\textwidth}
\begin{flushleft} \large
\emph{Author (s):}\\
Nathan \textsc{Hill}, C3334136\\
Jacob \textsc{Saunders}, C3262240\\
\end{flushleft}
\end{minipage}\\[1.5cm]

% DATE SECTION

\vfill % Fill the rest of the page with whitespace
{\large \today} \\ [1.5cm] % Adjusted spacing here
\end{titlepage}

%----------------------------------------------------------------------------------------
%	Document Contents
%----------------------------------------------------------------------------------------

\section*{Subject area}:

\section*{Data set}:

\section*{Topics covered}:

\subsubsection*{ -Topic 1- }

\subsubsection*{ -Topic 2- }

\subsubsection*{ -Topic 3- }

\subsubsection*{ -Topic 4- }

%----------------------------------------------------------------------------------------
%	Bibliography
%----------------------------------------------------------------------------------------

\bibliographystyle{elsarticle-num}
\bibliography{assets/_biblo.bib}

\end{document}